Improving reproducibility of scientific software using Nix/NixOS: A case study on the preCICE ecosystem

### 2024-05-15 paper/paper.tex
feat(paper): Add tutorials subsection

In this paper, we analyze how well Nix and NixOS can improve reproducibility of scientific software.
To this end, we study the preCICE ecosystem~\cite{preCICEv2} as an example.
preCICE is a coupling library for partitioned multiphysics simulations.
The ecosystem includes diverse legacy solvers, adapters, language bindings, and testcases besides the coupling library itself making it a challenging and representative testcase.
We try to build the complete ecosystem using Nix and run all testcases.

@@

preCICE provides bindings\footnote{\url{https://github.com/precice/fortran-module}} to the fortran programming language.
We were able to package these by providing custom phases for the \texttt{buildPhase} and the \texttt{installPhase}.
The resulting binary could run the example that comes with the module.

@@

\begin{sloppypar}
Most of the functionality that can be found in the official VM is also available in our NixOS VM and all preCICE tutorials could be run successfully, except \texttt{turek-hron-fsi2} as the tutorial needs \texttt{swak4foam}, an add-on to OpenFOAM, we were not able to package, \texttt{flow-over-heated-plate-steady-state} as the manual steps involved were unclear, \texttt{partitioned-heat-conduction-complex} and \texttt{partitioned-heat-conduction-direct} as the \texttt{mshr} python dependency was missing.
\end{sloppypar}

>>>>

In this paper, we analyze how well Nix and NixOS can improve reproducibility of scientific software.
To this end, we study the preCICE ecosystem~\cite{preCICEv2} as an example.
preCICE is a coupling library for partitioned multiphysics simulations.
The ecosystem includes diverse legacy solvers, adapters, language bindings, and tutorials besides the coupling library itself making it a challenging and representative testcase.
We try to build the complete ecosystem using Nix and run all tutorials.

@@

preCICE provides bindings\footnote{\url{https://github.com/precice/fortran-module}} to the Fortran programming language.
We were able to package these by providing custom phases for the \texttt{buildPhase} and the \texttt{installPhase}.
The resulting binary could run the example that comes with the module.

@@

\subsection{preCICE tutorials}

The preCICE tutorials included in the distribution can not only be used as reference examples on how to couple different adapters, but also to verify functionality of adapter and solver installations by executing them.

All v202211.0 test can be run successfully inside our NixOS VM and also in a \texttt{nix develop} environment, except \texttt{turek-hron-fsi2} and \texttt{flow-over-heated-plate-steady-state}.

We could not manage to build \texttt{swak4foam}, an add-on to OpenFOAM that is needed in the \texttt{turek-hron-fsi2} tutorial, from source.
The official preCICE VM uses a prebuilt version of swak4foam, however also patching the prebuilt binary did not work.

The second simulation we were not able to run is \texttt{flow-over-heated-plate-steady-state}.
This testcase needs code-aster to work, that we managed to package as already mentioned.
However, when running the simulation after executing the manual steps as written in the documentation, there occurs a segmentation fault, crashing code-aster when the python code tries to call a Fortran subroutine.
